---
title: "mississippi-data-assembly-Oct"
author: "Kate Morkeski"
date: "`r Sys.Date()`"
output: html_document
---

```{r}

library(here)
library(readxl)
library(tidyverse)
library(geosphere)
library(lubridate)
library(stats)

```

# Read in CTD and bottle data

```{r}

compiled <- read_excel(here('MSR_2_Compiled.xlsx')) 

```

# Format date and time

```{r}

compiled <- compiled %>%
  rename(time_utc = "Time (UTC)") %>%
  rename(station = Station_ID) %>%
  rename(cast = Cast_number) %>%
  rename(niskin = Niskin_ID) %>%
  rename(lat = "Latitude [decimal degrees]") %>%
  rename(long = "Longitude [decimal degrees]") %>%
  rename(temperature = "CTD_Potential_Temp [deg_C]") %>%
  rename(depth_m = "Depth [m]") %>%
  mutate(long = long*-1)

# combine date and time
compiled$Date <- as.Date(compiled$Date, format = "%B %d %Y") 
compiled$time_utc <- as.character(compiled$time_utc)
compiled$time_utc <- gsub("1899-12-31 ", "", compiled$time_utc)
compiled$date_time_utc <- paste(compiled$Date, compiled$time_utc)
compiled$date_time_utc <- as.POSIXct(compiled$date_time_utc, tz = "UTC", format="%Y-%m-%d %H:%M:%OS")

compiled <- compiled %>% relocate(date_time_utc, .before = station)

```
# Create map

```{r}

# very generic state boundary
# need more detailed map data
us_map <- map_data("state")

louisiana <- us_map %>% 
  filter(region %in% "louisiana") 

mouth <- louisiana %>% filter(long > -91.5 & lat < 30.5)

mapping <- compiled %>% 
  select(date_time_utc, station, cast, niskin, lat, long, temperature, depth_m, CTDSAL_PSS78) 

ggplot(mapping, aes(x = long, y = lat)) +
  geom_point() +
  theme_minimal()

ggplot(louisiana, aes(x = long, y = lat)) +
  geom_path() +
  geom_point(data = mapping, aes(x = long, y = lat, color = CTDSAL_PSS78)) +  
  theme_minimal()

ggplot(louisiana, aes(x = long, y = lat)) +
  geom_path() +
  geom_point(data = mapping, aes(x = long, y = lat, color = station)) +  
  theme_minimal()  

ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_point(data = mapping, aes(x = long, y = lat, color = CTDSAL_PSS78)) +  
  theme_minimal()

ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_text(data = mapping, aes(x = long, y = lat, label = station)) 
ggsave('Oct_station_map_all.png')

ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_text(data = mapping, aes(x = long, y = lat, label = station), size = 3) +  
  coord_fixed(xlim = c(-90.0, -89.25), ylim = c(28.5, 29.25))
ggsave('Oct_station_map_plume.png')

ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_path(data = mapping, aes(x = long, y = lat, color = CTDSAL_PSS78)) +  
  coord_fixed(xlim = c(-90.0, -89.25), ylim = c(28.5, 29.25))+
  theme_minimal()

```

# Towards join with GO data

```{r}

# read in GO data
GOfiles <- dir(here("MSR_GO_data_Oct/"), "*.txt") # get file names

GO_Oct <- GOfiles %>% map_dfr(~ read_tsv(here("MSR_GO_data_Oct/", .), show_col_types = FALSE, ))

```

```{r}

# handle date and time columns
GO_Oct <- rename(GO_Oct, PC_Date = "PC Date")
GO_Oct <- rename(GO_Oct, PC_Time = "PC Time")

# combine date and time
GO_Oct$PC_Date <- as.character(GO_Oct$PC_Date) 
GO_Oct$PC_Time <- as.character(GO_Oct$PC_Time) 
GO_Oct$date_time_local <- paste(GO_Oct$PC_Date, GO_Oct$PC_Time)
GO_Oct$date_time_local <- as.POSIXct(GO_Oct$date_time_local, tz = "America/Chicago", format="%d/%m/%y %H:%M:%OS")
GO_Oct$date_time_utc <- with_tz(GO_Oct$date_time_local, "UTC")

ggplot(GO_Oct, aes(x = date_time_local, y = "equ press")) +
  geom_point() +
  theme_minimal()

ggplot(GO_Oct, aes(x = date_time_utc, y = "equ press")) +
  geom_point() +
  theme_minimal()

```
# Identify surface stations

```{r}

# identify surface rows
mapping <-mapping %>% 
  mutate(surface = case_when(station < 16 & depth_m < 1 ~ "Y",
                             station > 15 & depth_m < 2 ~ "Y",
                             TRUE ~ "N"
  )) 

surface_bottles <-mapping %>% filter(surface == "Y")

surface_bottles <- surface_bottles %>% select(-surface)

```
# Match bottle data by time

```{r}

# match bottle data to GO data based on datetime
# get lat, long, T, S from bottle data

for (i in 1:nrow(surface_bottles)) {
  
  surface_bottles$GO_row[i] <- which.min(abs(GO_Oct$date_time_utc - surface_bottles$date_time_utc[i]))
  
}

# for surface bottles, average temperature, depth, and salinity by station and cast
# retain lat, long, and GO_row
surface_avgs <- surface_bottles %>% 
  group_by(station, cast) %>%
  summarize(lat = mean(lat),
            long = mean(long), 
            temperature = mean(temperature),
            depth_m = mean(depth_m),
            CTDSAL_PSS78 = mean(CTDSAL_PSS78),
            GO_row = mean(GO_row))

# round GO_row to integer
surface_avgs$GO_row <- round(surface_avgs$GO_row, 0)

GO_Oct$GO_row <- seq.int(nrow(GO_Oct))

GOplus <- left_join(GO_Oct, surface_avgs, by = "GO_row")

# check which bottle rows are used
ctd_check <- GOplus %>% 
  filter(!is.na(station))

GOplus <- GOplus %>%
  select(-GO_row, -station, -cast, -depth_m)

```
# Interpolate CTD data to GO timestamps

```{r}

# # interpolate atmospheric pressure to GO data time stamps
# GO_Patm <- data.frame(approx(Oct_Patm$datetime, Oct_Patm$Patm_mbar_avg, xout = GO_out$date_time_utc, rule = 2, method = "linear"))
# GO_out <- cbind(GO_out, GO_Patm)
# GO_out <- GO_out %>% rename(Patm_mbar = y) %>% select(-x) %>% mutate(Patm_atm = Patm_mbar/1013.25)

```


# format GO data

```{r}

GO_out <- GOplus %>% 
  filter(Type != "SHUT DOWN" ) 

```

# Read in met data

```{r}

#https://mesonet.agron.iastate.edu/request/download.phtml?network=LA_ASOS
# one site is MSY airport and the other Is a rig/platform south of the estuary (named DSF). 

MSY_met <- read.csv("MSY_Met.csv")
rig_met <- read.csv("MSR_Met_Data.csv")

MSY_met <- rename(MSY_met, datetime = valid)

# set time zone 
MSY_met$datetime <- strptime(MSY_met$datetime,"%m/%d/%Y %R")
rig_met$datetime <- strptime(rig_met$datetime,"%m/%d/%Y %R")
#tz(MSY_met$datetime) <- "America/Chicago"
#tz(rig_met$datetime) <- "America/Chicago"
tz(MSY_met$datetime) <- "UTC"
tz(rig_met$datetime) <- "UTC"

rig_met$alti <- as.numeric(rig_met$alti)
MSY_met$mslp <- as.numeric(MSY_met$mslp)

GO_out <- rename(GO_out, equ_press = "equ press")
GO_out$equ_press <- as.numeric(GO_out$equ_press)
GO_out <- rename(GO_out, licor_press = "licor press")
GO_out$licor_press <- as.numeric(GO_out$licor_press)
GO_out$equ_calc <- GO_out$equ_press + GO_out$licor_press

# need POSIXct for graphing together
MSY_met$datetime <- as.POSIXct(MSY_met$datetime)
rig_met$datetime <- as.POSIXct(rig_met$datetime)

# format wind speed for plotting
MSY_met <- rename(MSY_met, speed_mph = "speed..mph.")
MSY_met$speed_mph <- replace(MSY_met$speed_mph, MSY_met$speed_mph == "M", NA)
MSY_met$speed_mph <- as.numeric(MSY_met$speed_mph)

rig_met <- rename(rig_met, speed_mph = "speed..mph.")
rig_met$speed_mph <- replace(rig_met$speed_mph, rig_met$speed_mph == "M", NA)
rig_met$speed_mph <- as.numeric(rig_met$speed_mph)

# isolate dates of Oct cruise
Oct_MSY_met <- MSY_met %>% filter(month(datetime) == 10)
Oct_MSY_met <- Oct_MSY_met %>% filter(day(datetime) > 11 & day(datetime) < 14 ) 
Oct_rig_met <- rig_met %>% filter(month(datetime) == 10)
Oct_rig_met <- Oct_rig_met %>% filter(day(datetime) > 11 & day(datetime) < 14 ) 

# create data frame to plot met stations relative to cruise stations
station <- c(MSY_met$station[1], rig_met$station[1])
long = c(MSY_met$lon[1], rig_met$lon[1])
lat = c(MSY_met$lat[1], rig_met$lat[1])
met_stns <- data.frame(station, long, lat)

```

# plot met data

```{r}

# plot 2022 pressure data from both sites
ggplot(MSY_met, aes(x = datetime, y = alti, color = "MSY")) +
  geom_point() +
  geom_point(data = rig_met, aes(x = datetime, y = alti, color = "rig")) + 
  theme_minimal()

# plot 2022 wind data from both sites
ggplot(MSY_met, aes(x = datetime, y = speed_mph, color = "MSY")) +
  geom_point() +
  geom_point(data = rig_met, aes(x = datetime, y = speed_mph, color = "rig")) + 
  theme_minimal()

# plot pressure data from Oct cruise dates
# alti
ggplot(Oct_MSY_met, aes(x = datetime, y = alti, color = "MSY")) +
  geom_point() +
  geom_point(data = Oct_rig_met, aes(x = datetime, y = alti, color = "rig")) + 
  theme_minimal()
# mslp
ggplot(Oct_MSY_met, aes(x = datetime, y = mslp, color = "MSY")) +
  geom_point() +
  theme_minimal()
# mslp and equ pressure
scale = 1
# local time
ggplot(Oct_MSY_met, aes(x = datetime, y = mslp, color = "MSY")) +
  geom_point() +
  geom_point(data = GO_out, aes(x = date_time_local, y = equ_calc, color = "equ")) +
  scale_y_continuous(limits = c(1012, 1024), sec.axis = sec_axis(~.*scale, name="mslp")) +
  theme_minimal()

# utc
ggplot(Oct_MSY_met, aes(x = datetime, y = mslp, color = "MSY")) +
  geom_point() +
  geom_point(data = GO_out, aes(x = date_time_utc, y = equ_calc, color = "equ")) +
  scale_y_continuous(limits = c(1012, 1024), sec.axis = sec_axis(~.*scale, name="mslp")) +
  theme_minimal()

# plot locations of met stations
ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_text(data = mapping, aes(x = long, y = lat, label = station)) +  
  geom_text(data = met_stns, aes(x = long, y = lat, label = station))+
  theme_minimal()  

```

# Convert alti from inches Hg to mbar
```{r}

# multiply alti by 33.8639 to convert from inches Hg to mbar
Oct_MSY_met <- Oct_MSY_met %>%
  mutate(Patm_mbar = alti*33.8639)

Oct_rig_met <- Oct_rig_met %>%
  mutate(Patm_mbar = alti*33.8639)

# plot pressure data 
ggplot(Oct_MSY_met, aes(x = datetime, y = Patm_mbar, color = "MSY")) +
  geom_point() + 
  geom_point(data = Oct_rig_met, aes(x = datetime, y = Patm_mbar, color = "rig")) + 
  theme_minimal()

```

# Interpolate met data and average across MSY and DSF stations

```{r}

# Oct
Oct_MSY_Patm <- data.frame(approx(Oct_MSY_met$datetime, Oct_MSY_met$Patm_mbar, xout = Oct_rig_met$datetime, rule = 2, method = "linear"))

Oct_rig_Patm <- Oct_rig_met %>% select(datetime, Patm_mbar)

Oct_Patm <- cbind(Oct_MSY_Patm, Oct_rig_Patm)
Oct_Patm <- Oct_Patm %>% 
  rename(Patm_mbar_MSY = y) %>% 
  rename(Patm_mbar_rig = Patm_mbar) 

Oct_Patm <- Oct_Patm %>%
  mutate(Patm_mbar_avg = (Patm_mbar_MSY + Patm_mbar_rig)/2)

ggplot(Oct_Patm, aes(x = datetime, y = Patm_mbar_MSY, color = "MSY")) +
  geom_point() + 
  geom_point(data = Oct_Patm, aes(x = datetime, y = Patm_mbar_rig, color = "rig")) + 
  geom_point(data = Oct_Patm, aes(x = datetime, y = Patm_mbar_avg, color = "avg")) + 
  theme_minimal()

ggplot(Oct_MSY_met, aes(x = datetime, y = Patm_mbar, color = "MSY")) +
  geom_point() + 
  geom_point(data = Oct_rig_met, aes(x = datetime, y = Patm_mbar, color = "rig")) + 
  theme_minimal()

```

```{r}

# interpolate atmospheric pressure to GO data time stamps
GO_Patm <- data.frame(approx(Oct_Patm$datetime, Oct_Patm$Patm_mbar_avg, xout = GO_out$date_time_utc, rule = 2, method = "linear"))
GO_out <- cbind(GO_out, GO_Patm)
GO_out <- GO_out %>% rename(Patm_mbar = y) %>% select(-x) %>% mutate(Patm_atm = Patm_mbar/1013.25)

# interpolate wind data to GO time series
# use only MSY because don't have real height of DSF sensor
GO_wind <- data.frame(approx(Oct_MSY_met$datetime, Oct_MSY_met$speed_mph, xout = GO_out$date_time_utc, rule = 2, method = "linear"))
GO_out <- cbind(GO_out, GO_wind)
GO_out <- GO_out %>% rename(speed_mph_MSY = y) %>% select(-x) %>% mutate(speed_m_s_MSY = speed_mph_MSY*0.44704)

# MSY temperature and dewpoint data have better resolution than DSF, so use only MSY 
# interpolate wind data to GO time series
GO_temp <- data.frame(approx(Oct_MSY_met$datetime, Oct_MSY_met$tmpc, xout = GO_out$date_time_utc, rule = 2, method = "linear"))
GO_out <- cbind(GO_out, GO_temp)
GO_out <- GO_out %>% rename(temp_air_C = y) %>% select(-x)
GO_dwpc <- data.frame(approx(Oct_MSY_met$datetime, Oct_MSY_met$dwpc, xout = GO_out$date_time_utc, rule = 2, method = "linear"))
GO_out <- cbind(GO_out, GO_dwpc)
GO_out <- GO_out %>% rename(dewpoint_C = y) %>% select(-x)
# calculate vapor pressure: https://www.weather.gov/media/epz/wxcalc/vaporPressure.pdf
GO_out <- GO_out %>%
  mutate(pWater_mbar = 6.11*10^((7.5*dewpoint_C)/(237.3+dewpoint_C))) %>%
  mutate(pWater_atm = pWater_mbar/1013.25)         

```

# plot met data

```{r}

ggplot(Oct_MSY_met, aes(x = datetime, y = tmpc, color = "MSY")) +
  geom_point() + 
  geom_point(data = Oct_rig_met, aes(x = datetime, y = tmpc, color = "rig")) + 
  theme_minimal()

ggplot(Oct_MSY_met, aes(x = datetime, y = dwpc, color = "MSY")) +
  geom_point() + 
  geom_point(data = Oct_rig_met, aes(x = datetime, y = dwpc, color = "rig")) + 
  theme_minimal()

ggplot(GO_out, aes(x = date_time_utc, y = speed_m_s_MSY)) +
  geom_point() + 
  labs(x = "datetime", y = "wind speed m/s") +
  theme_minimal()

ggplot(GO_out, aes(x = date_time_utc, y = temp_air_C, color = "air temperature")) +
  geom_point() + 
  geom_point(data = GO_out, aes(x = date_time_utc, y = dewpoint_C, color = "dewpoint")) + 
  labs(x = "datetime", y = "Temperature C") +
  theme_minimal()

ggplot(GO_out, aes(x = date_time_utc, y = Patm_atm, color = "P total")) +
  geom_point() + 
  labs(x = "datetime", y = "Pressure (atm)") +
  theme_minimal()

ggplot(GO_out, aes(x = date_time_utc, y = pWater_atm, color = "P H20")) +
  geom_point() + 
  labs(x = "datetime", y = "Pressure (atm)") +
  theme_minimal()

```

# Write file for input into pCO2sys

```{r}

GO_out <- GO_out %>% 
  select(-speed_mph_MSY, -dewpoint_C)

write.csv(GO_out, here('Oct_GO_CTD_met.csv'), row.names = FALSE)

```

## October cruise

# Pro Oceanus pCO2
```{r}

# read Pro Oceanus pCO2 data
POceanus <- read.csv(here('20221012pco2.csv')) 

# delete first row
POceanus <- POceanus[-(1),]
POceanus <- POceanus[-(69135:69147),]

# delete rows before underway water flowing
POceanus <- POceanus[-(1:5351),] #noon

```
# combine date time info and set column types

```{r}

POceanus$date_time <- paste(POceanus$Year, "-" ,POceanus$Month, "-", POceanus$Day, "x", POceanus$Hour, ":", POceanus$Minute, ":", POceanus$Second)
POceanus <- POceanus %>% relocate(date_time, .before = "Zero.A.D")

POceanus$date_time <- gsub(" ", "", POceanus$date_time)
POceanus$date_time <- gsub("x", " ", POceanus$date_time)

POceanus$date_time <- as.POSIXct(POceanus$date_time, tz = "America/Chicago", format="%Y-%m-%d %H:%M:%OS")

POceanus$CO2 <- as.double(POceanus$CO2)
POceanus$Humidity <- as.double(POceanus$Humidity)
POceanus$Cell.gas.pressure <- as.double(POceanus$Cell.gas.pressure)
POceanus$IRGA.temperature <- as.double(POceanus$IRGA.temperature)
POceanus$Day <- as.double(POceanus$Day)

```


# Plot Pro Oceanus data

```{r}

ggplot(POceanus, aes(x = date_time, y = CO2))+
  geom_point()+
  theme_minimal()

ggplot(POceanus, aes(x = date_time, y = Humidity))+
  geom_point()+
  theme_minimal()

ggplot(POceanus, aes(x = date_time, y = Cell.gas.pressure))+
  geom_point()+
  theme_minimal()

scale = 0.2    
ggplot(POceanus, aes(x = date_time, y = CO2)) +
  geom_point(aes(color = "pCO2 ppm")) +
  geom_point(aes(y = Cell.gas.pressure/scale, color = "Cell gas pressure")) +
  scale_y_continuous(sec.axis = sec_axis(~.*scale, name="pressure")) +
  labs(x = "Date and Time", y = "pCO2 ppm", color = "") +
  scale_color_manual(values = c("orange2", "gray30"))+
  theme_minimal()


ggplot(POceanus, aes(x = date_time, y = IRGA.temperature))+
  geom_point()+
  theme_minimal()

scale = 0.005
ggplot(POceanus, aes(x = date_time, y = CO2)) +
  geom_point(aes(color = "pCO2 ppm")) +
  geom_point(aes(y = Humidity/scale, color = "Humidity")) +
  #scale_x_continuous(breaks = seq(0, 336, 24)) +
  scale_y_continuous(sec.axis = sec_axis(~.*scale, name="Humidity (units)")) +
  labs(x = "Date and Time", y = "pCO2 ppm", color = "") +
  scale_color_manual(values = c("orange2", "gray30"))+
  theme_minimal()


```
# data cleaning steps

```{r}

outliers <- POceanus %>% filter(CO2 > 4000)

# two outliers with CO2 > 4000, 
# second at 2022-10-12 06:57:16 when instrument was started for the day # filter this one out
# first at 2022-10-11 19:29:13
# was this a zero? 
# after this, CO2 and pressure drop, and take a while to stabilize again (circa 1 hr)
# CO2 seems to return to where it was

# take a closer look at October 11
POceanus_11 <- POceanus %>% filter(Day < 12)
ggplot(POceanus_11, aes(x = date_time, y = CO2))+
  geom_point()+
  theme_minimal()

scale = 0.05    
ggplot(POceanus_11, aes(x = date_time, y = CO2)) +
  geom_point(aes(color = "pCO2 ppm")) +
  geom_point(aes(y = (Cell.gas.pressure-800)/scale, color = "Cell gas pressure")) +
  scale_y_continuous(sec.axis = sec_axis(~800+.*scale, breaks = seq(800, 1300, 50), name="pressure")) +
  labs(x = "Date and Time", y = "pCO2 ppm", color = "") +
  scale_color_manual(values = c("orange2", "gray30"))+
  theme_minimal()

# filter out points above 4k
POceanus_11_under4k <- POceanus_11 %>% filter(CO2 < 4000)
# rename columns to match GO data frame
POceanus_11_under4k <- rename(POceanus_11_under4k, date_time_gmt = date_time)
POceanus_11_under4k <- rename(POceanus_11_under4k, CO2_ppm = CO2)

# plot data below 4k
ggplot(POceanus_11_under4k, aes(x = date_time_gmt, y = CO2_ppm))+
  geom_point()+
  theme_minimal()

# filter out points above 4k
POceanus_under4k <- POceanus %>% filter(CO2 < 4000)
# rename columns to match GO data frame
POceanus_under4k <- rename(POceanus_under4k, date_time_gmt = date_time)
POceanus_under4k <- rename(POceanus_under4k, CO2_ppm = CO2)

# plot data below 4k
ggplot(POceanus_under4k, aes(x = date_time_gmt, y = CO2_ppm))+
  geom_point()+
  theme_minimal()

```

#clean up long equilibration period after 2022-10-11 19:29:13

```{r}

POceanus_under4k$date_time_num <- as.numeric(POceanus_under4k$date_time_gmt)

POceanus_junk <- POceanus_under4k %>% filter(Day == 11 & Hour == "19" & Minute == "28")
start_junk <- POceanus_junk$date_time_num[1]

POceanus_junk <- POceanus_under4k %>% filter(Day == 11 & Hour == "19" & Minute == "50")
end_junk <- POceanus_junk$date_time_num[1]

POceanus_under4k <- POceanus_under4k %>%
  mutate(leg = case_when(start_junk < date_time_num & date_time_num < end_junk ~ "junk"))

POceanus_junk <- POceanus_under4k %>% filter(Day == 11 & Hour == "13" & Minute == "52")
end_canal <- POceanus_junk$date_time_num[1]

POceanus_junk <- POceanus_under4k %>% filter(Day == 11 & Hour == "22" & Minute == "00")
start_night <- POceanus_junk$date_time_num[1]
POceanus_junk <- POceanus_under4k %>% filter(Day == 12 & Hour == "07" & Minute == "30")
end_night <- POceanus_junk$date_time_num[1]

POceanus_night <- POceanus_under4k %>% filter(start_night < date_time_num & date_time_num < end_night)
end_night - start_night

POceanus_under4k <- POceanus_under4k %>%
  mutate(leg = case_when(start_junk < date_time_num & date_time_num < end_junk ~ "junk",
                         date_time_num < end_canal ~ "junk",
                         start_night < date_time_num & date_time_num < end_night ~ "junk"))

ggplot(POceanus_under4k, aes(x = date_time_gmt, y = CO2_ppm, color = leg))+
  geom_point()+
  theme_minimal()

POceanus_clean <- POceanus_under4k %>% filter(is.na(leg))

POceanus_clean <- POceanus_clean %>% select(-date_time_num, -leg)

ggplot(POceanus_clean, aes(x = date_time_gmt, y = CO2_ppm))+
  geom_point()+
  theme_minimal()

POceanus_clean <- POceanus_clean %>% 
  rename(date_time_local = date_time_gmt) 

```

# 

```{r}


GO_out <- GO_out %>% rename(CO2_ppm = "CO2 um/m")


```

# Plot raw GO data

```{r}

ggplot(GO_out, aes(x = date_time_local, y = CO2_ppm, color = "GO"))+
  geom_point()+
  geom_point(data = POceanus_clean, aes(x = date_time_local, y = CO2_ppm, color = "Pro Oceanus")) + 
  labs(x = "datetime", y = "CO2 ppm") +
  theme_minimal()

ggplot(GO_out, aes(x = date_time_utc, y = CO2_ppm, color = "Type"))+
  geom_point()+
  theme_minimal()

```

```{r}

GO_Oct_EQU <- GO_out %>% filter(Type == "EQU" | Type == "EQU-DRAIN")

ggplot(GO_Oct_EQU, aes(x = date_time_utc, y = CO2_ppm))+
  geom_point()+
  theme_minimal()

# clean up beginning and end of Oct 11
GO_junk <- GO_out %>% filter(date_time_local < "2022-10-11 14:47:27")
GO_junk <- GO_out %>% filter(Type == "EQU" & date_time_utc < "2022-10-11 14:47:27")
GO_out <- GO_out %>%
  mutate(leg = case_when(Type == "EQU" & date_time_utc < "2022-10-11 14:47:27" ~ "junk",
                         Type == "EQU-DRAIN" & date_time_utc < "2022-10-11 14:47:27" ~ "junk",
                         Type =="EQU" & date_time_utc > "2022-10-11 23:00:00" & date_time_utc < "2022-10-12 01:00:00" ~ "junk",
                         Type =="EQU-DRAIN" & date_time_utc > "2022-10-11 23:00:00" & date_time_utc < "2022-10-12 01:00:00" ~ "junk"))

ggplot(GO_out, aes(x = date_time_utc, y = CO2_ppm, color = leg))+
  geom_point()+
  theme_minimal()

GO_clean <- GO_out %>% filter(is.na(leg))

GO_Oct_EQU <- GO_clean %>% filter(Type == "EQU" | Type == "EQU-DRAIN")

ggplot(GO_Oct_EQU, aes(x = date_time_utc, y = CO2_ppm))+
  geom_point()+
  theme_minimal()

```

```{r}

# GO_Oct_ATM <- GO_Oct %>% filter(Type == "ATM" | Type == "ATM-DRAIN")
# 
# ggplot(GO_Oct_ATM, aes(x = date_time_gmt, y = CO2_ppm))+
#   geom_point()+
#   theme_minimal()
# 
# GO_Oct_STD <- GO_Oct %>% 
#   filter(Type == "STD1" | Type == "STD1-DRAIN" | Type == "STD2" | Type == "STD2-DRAIN" | Type == "STD3" | Type == "STD3-DRAIN" | Type == "STD4" | Type == "STD4-DRAIN" )
# 
# ggplot(GO_Oct_STD, aes(x = date_time_gmt, y = CO2_ppm))+
#   geom_point()+
#   theme_minimal()

```

```{r}

ggplot(GO_Oct_EQU, aes(x= date_time_local, y = CO2_ppm, color = "General Oceanics"))+ geom_point() +geom_point(data = POceanus_clean, aes(color = "Pro Oceanus"))

write.csv(POceanus_clean, here('Pro_Oceanus_pCO2_Mississippi_Oct_2022.csv'), row.names = FALSE)

```
# Read in GPS data

```{r}

gps_oct <- read_delim("sync_gps_ctd_oct.txt", show_col_types = FALSE)

gps_oct$dist_km <- gps_oct$Distancem/1000

gps_oct <- gps_oct %>% relocate(dist_km, .after = "Distancem")

tz(gps_oct$DateLocal) <- "America/Chicago"
tz(gps_oct$DateGMT) <- "UTC"

```

```{r}

ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_point(data = mapping, aes(x = long, y = lat)) +  
  theme_minimal()

ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_point(data = gps_oct, aes(x = Longitude, y = Latitude)) +  
  theme_minimal()

ggplot(GO_out, aes(x = date_time_local, y = lat)) +
  geom_point() +
  geom_point(data = gps_oct, aes(x = DateLocal, y = Latitude)) +  
  theme_minimal()

ggplot(GO_out, aes(x = date_time_local, y = lat, color = "GO")) +
  geom_point() +
  geom_point(data = gps_oct, aes(x = DateGMT, y = Latitude, color = "GPS")) +  
  theme_minimal()


```


```{r}
# match gps data to GO data based on datetime

for (i in 1:nrow(GO_out)) {
  
  GO_out$ind[i] <- which.min(abs(GO_out$date_time_local[i] - gps_oct$DateGMT))
  
}

gps_oct$row_num <- seq.int(nrow(gps_oct))
gps_oct$bottlerow <- as.character(gps_oct$row_num)
GO_out$bottlerow <- as.character(GO_out$ind)

GO_out <- left_join(GO_out, gps_oct, by = "bottlerow")

GO_out <- GO_out %>% 
  select(-HorizontalAccuracym, -Altitudem, -VerticalAccuracym, -Distancem, -AverageSpeedms, -Coursedeg, -TrueHeadingdeg, -MagneticHeadingdeg, -HeadingAccuracydeg, -GlideRatio, -HeartRatebpm, -VarName6, -CTDMDY, -CTDHMS, -CTDTELAPS, -row_num) 

GO_out <- GO_out %>% 
  select(-lat, -long, -equ_calc, -bottlerow, -DateLocal, -DateGMT, -Timesec, -CTDCOND, -CTDPRS, -CTDDENS)

GO_out <- GO_out %>% 
  filter(is.na(leg))

GO_out <- GO_out %>% 
  select(-leg, -ind)

```

```{r}

# compare T and S in GPS data to station data

ggplot(GO_out, aes(x = date_time_local, y = CTDSAL_PSS78, color = "station sal"))+
  geom_point()+
  geom_point(aes(x = date_time_local, y = CTDSAL, color = "interp sal")) + 
  labs(x = "datetime", y = "salinity") +
  theme_minimal()

ggplot(GO_out, aes(x = date_time_local, y = temperature, color = "station temp"))+
  geom_point()+
  geom_point(aes(x = date_time_local, y = CTDTMP, color = "interp temp")) + 
  labs(x = "datetime", y = "temperature C") +
  theme_minimal()

# underway temperature and salinity from GPS data set appear to be no good
GO_out <- GO_out %>% 
  select(-CTDTMP, -CTDSAL)
         
```

# interpolate station T and S 

```{r}

temp_interp <- data.frame(approx(ctd_check$date_time_utc, ctd_check$temperature, xout = GO_out$date_time_utc, rule = 2, method = "linear"))
GO_out <- cbind(GO_out, temp_interp)
GO_out <- GO_out %>% select(-x, -temperature) %>% rename(temperature = y)  

sal_interp <- data.frame(approx(ctd_check$date_time_utc, ctd_check$CTDSAL_PSS78, xout = GO_out$date_time_utc, rule = 2, method = "linear"))
GO_out <- cbind(GO_out, sal_interp)
GO_out <- GO_out %>% select(-x, -CTDSAL_PSS78) %>% rename(CTDSAL_PSS78 = y)  


```


# write file for pCO2sys

```{r}

write.csv(GO_out, here('GO_Mississippi_Oct_NaN.csv'), na = "NaN", row.names = FALSE)

write.csv(GO_out, here('GO_Mississippi_Oct_999.csv'), na = "-999", row.names = FALSE)

```






