---
title: "mississippi-data-assembly-Oct"
author: "Kate Morkeski"
date: "`r Sys.Date()`"
output: html_document
---

```{r}

library(here)
library(readxl)
library(tidyverse)
library(geosphere)
library(lubridate)
library(stats)

```

# Read in CTD and bottle data

```{r}
#TODO: change to Oct
#compiled <- read_excel(here('MSR_1_Compiled.xlsx')) 
#salinity <- read_excel(here('MSR_1_Salinity.xlsx'))
```

# Format date and time

```{r}

compiled <- compiled %>%
  rename(time_utc = "Time (UTC)") %>%
  rename(station = Station_ID) %>%
  rename(cast = Cast_number) %>%
  rename(niskin = Niskin_ID) %>%
  rename(lat = "Latitude [decimal degrees]") %>%
  rename(long = "Longitude [decimal degrees]") %>%
  rename(temperature = "CTD_Potential_Temp [deg_C]") %>%
  rename(depth_m = "Depth [m]")

# combine date and time
compiled$Date <- as.character(compiled$Date) 
compiled$time_utc <- as.character(compiled$time_utc)
compiled$time_utc <- gsub("1899-12-31 ", "", compiled$time_utc)
compiled$date_time_utc <- paste(compiled$Date, compiled$time_utc)
compiled$date_time_utc <- as.POSIXct(compiled$date_time_utc, tz = "UTC", format="%Y-%m-%d %H:%M:%OS")

compiled <- compiled %>% relocate(date_time_utc, .before = station)

```
# Create map

```{r}

# very generic state boundary
# need more detailed map data
us_map <- map_data("state")

louisiana <- us_map %>% 
  filter(region %in% "louisiana") 

mouth <- louisiana %>% filter(long > -91.5 & lat < 30.5)

mapping <- compiled %>% 
  select(date_time_utc, station, cast, niskin, lat, long, temperature, depth_m, CTDSAL_PSS78) 

ggplot(louisiana, aes(x = long, y = lat)) +
  geom_path() +
  geom_point(data = mapping, aes(x = long, y = lat, color = CTDSAL_PSS78)) +  
  theme_minimal()

ggplot(louisiana, aes(x = long, y = lat)) +
  geom_path() +
  geom_point(data = mapping, aes(x = long, y = lat, color = station)) +  
  theme_minimal()  

ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_point(data = mapping, aes(x = long, y = lat, color = CTDSAL_PSS78)) +  
  theme_minimal()

ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_text(data = mapping, aes(x = long, y = lat, label = station)) +  
  theme_minimal()

ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_text(data = mapping, aes(x = long, y = lat, label = station), size = 3) +  
  coord_fixed(xlim = c(-90.0, -89.25), ylim = c(28.5, 29.25))+
  theme_minimal()

ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_path(data = mapping, aes(x = long, y = lat, color = CTDSAL_PSS78)) +  
  coord_fixed(xlim = c(-90.0, -89.25), ylim = c(28.5, 29.25))+
  theme_minimal()

```

# Towards join with GO data

```{r}

#TODO: update to Oct
# read in GO data
#GOfiles <- dir(here("MSR_GO_data_May/"), "*.txt") # get file names

#GO_May <- GOfiles %>% map_dfr(~ read_tsv(here("MSR_GO_data_May/", .), show_col_types = FALSE, ))

```

```{r}

# handle date and time columns
GO_May <- rename(GO_May, PC_Date = "PC Date")
GO_May <- rename(GO_May, PC_Time = "PC Time")

# combine date and time
GO_May$PC_Date <- as.character(GO_May$PC_Date) 
GO_May$PC_Time <- as.character(GO_May$PC_Time) 
GO_May$date_time_utc <- paste(GO_May$PC_Date, GO_May$PC_Time)
GO_May$date_time_utc <- as.POSIXct(GO_May$date_time_utc, tz = "UTC", format="%d/%m/%y %H:%M:%OS")

```

# Match bottle data by time

```{r}

# match bottle data to GO data based on datetime
# get lat, long, T, S from bottle data

for (i in 1:nrow(surface_bottles)) {
  
  surface_bottles$GO_row[i] <- which.min(abs(GO_May$date_time_utc - surface_bottles$date_time_utc[i]))
  
}

# for surface bottles, average temperature, depth, and salinity by station and cast
# retain lat, long, and GO_row
surface_avgs <- surface_bottles %>% 
  group_by(station, cast) %>%
  summarize(lat = mean(lat),
            long = mean(long), 
            temperature = mean(temperature),
            depth_m = mean(depth_m),
            CTDSAL_PSS78 = mean(CTDSAL_PSS78),
            GO_row = mean(GO_row))

# round GO_row to integer
surface_avgs$GO_row <- round(surface_avgs$GO_row, 0)

GO_May$GO_row <- seq.int(nrow(GO_May))

GOplus <- left_join(GO_May, surface_avgs, by = "GO_row")

# check which bottle rows are used
ctd_check <- GOplus %>% 
  filter(!is.na(station))

```

# format GO data

```{r}

GO_out <- GO_out %>% 
  filter(Type != "SHUT DOWN") 

may_fug[may_fug=="-999"]<-NA
may_fug[may_fug=="NaN"]<-NA

```

# Read in met data

```{r}

GO_out <- read_csv('GO_Mississippi_May.csv')

#https://mesonet.agron.iastate.edu/request/download.phtml?network=LA_ASOS
# one site is MSY airport and the other Is a rig/platform south of the estuary (named DSF). 

MSY_met <- read.csv("MSY_Met.csv")
rig_met <- read.csv("MSR_Met_Data.csv")

MSY_met <- rename(MSY_met, datetime = valid)

# set time zone to GMT 
MSY_met$datetime <- strptime(MSY_met$datetime,"%m/%d/%Y %R")
rig_met$datetime <- strptime(rig_met$datetime,"%m/%d/%Y %R")
tz(MSY_met$datetime) <- "America/New_York"
tz(rig_met$datetime) <- "America/New_York"

rig_met$alti <- as.numeric(rig_met$alti)
MSY_met$mslp <- as.numeric(MSY_met$mslp)

GO_out <- rename(GO_out, equ_press = "equ press")
GO_out$equ_press <- as.numeric(GO_out$equ_press)
GO_out <- rename(GO_out, licor_press = "licor press")
GO_out$licor_press <- as.numeric(GO_out$licor_press)
GO_out$equ_calc <- GO_out$equ_press + GO_out$licor_press

# need POSIXct for graphing together
MSY_met$datetime <- as.POSIXct(MSY_met$datetime)
rig_met$datetime <- as.POSIXct(rig_met$datetime)

# format wind speed for plotting
MSY_met <- rename(MSY_met, speed_mph = "speed..mph.")
MSY_met$speed_mph <- replace(MSY_met$speed_mph, MSY_met$speed_mph == "M", NA)
MSY_met$speed_mph <- as.numeric(MSY_met$speed_mph)

rig_met <- rename(rig_met, speed_mph = "speed..mph.")
rig_met$speed_mph <- replace(rig_met$speed_mph, rig_met$speed_mph == "M", NA)
rig_met$speed_mph <- as.numeric(rig_met$speed_mph)

# isolate dates of Oct cruise
Oct_MSY_met <- MSY_met %>% filter(month(datetime) == 10)
Oct_MSY_met <- Oct_MSY_met %>% filter(day(datetime) > 11 & day(datetime) < 14 ) 
Oct_rig_met <- rig_met %>% filter(month(datetime) == 10)
Oct_rig_met <- Oct_rig_met %>% filter(day(datetime) > 11 & day(datetime) < 14 ) 

# create data frame to plot met stations relative to cruise stations
station <- c(MSY_met$station[1], rig_met$station[1])
long = c(MSY_met$lon[1], rig_met$lon[1])
lat = c(MSY_met$lat[1], rig_met$lat[1])
met_stns <- data.frame(station, long, lat)

```

# plot met data

```{r}

# plot 2022 pressure data from both sites
ggplot(MSY_met, aes(x = datetime, y = alti, color = "MSY")) +
  geom_point() +
  geom_point(data = rig_met, aes(x = datetime, y = alti, color = "rig")) + 
  theme_minimal()

# plot 2022 wind data from both sites
ggplot(MSY_met, aes(x = datetime, y = speed_mph, color = "MSY")) +
  geom_point() +
  geom_point(data = rig_met, aes(x = datetime, y = speed_mph, color = "rig")) + 
  theme_minimal()

# plot pressure data from Oct cruise dates
# alti
ggplot(Oct_MSY_met, aes(x = datetime, y = alti, color = "MSY")) +
  geom_point() +
  geom_point(data = Oct_rig_met, aes(x = datetime, y = alti, color = "rig")) + 
  theme_minimal()
# mslp
ggplot(Oct_MSY_met, aes(x = datetime, y = mslp, color = "MSY")) +
  geom_point() +
  theme_minimal()
# mslp and equ pressure
scale = 1
ggplot(Oct_MSY_met, aes(x = datetime, y = mslp, color = "MSY")) +
  geom_point() +
  geom_point(data = GO_out, aes(x = date_time_utc, y = equ_calc, color = "equ")) +
  scale_y_continuous(limits = c(1012, 1024), sec.axis = sec_axis(~.*scale, name="mslp")) +
  theme_minimal()

# plot locations of met stations
ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_text(data = mapping, aes(x = long, y = lat, label = station)) +  
  geom_text(data = met_stns, aes(x = long, y = lat, label = station))+
  theme_minimal()  

```

# Convert alti from inches Hg to mbar
```{r}

# multiply alti by 33.8639 to convert from inches Hg to mbar
Oct_MSY_met <- Oct_MSY_met %>%
  mutate(Patm = alti*33.8639)

Oct_rig_met <- Oct_rig_met %>%
  mutate(Patm = alti*33.8639)

# plot pressure data 
ggplot(Oct_MSY_met, aes(x = datetime, y = Patm, color = "MSY")) +
  geom_point() + 
  geom_point(data = Oct_rig_met, aes(x = datetime, y = Patm, color = "rig")) + 
  theme_minimal()

```

# Interpolate met data and average across MSY and DSF stations

```{r}

# Oct
Oct_MSY_Patm <- data.frame(approx(Oct_MSY_met$datetime, Oct_MSY_met$Patm, xout = Oct_rig_met$datetime, rule = 2, method = "linear"))

Oct_rig_Patm <- Oct_rig_met %>% select(datetime, Patm)

Oct_Patm <- cbind(Oct_MSY_Patm, Oct_rig_Patm)
Oct_Patm <- Oct_Patm %>% 
  rename(Patm_MSY = y) %>% 
  rename(Patm_rig = Patm) 

Oct_Patm <- Oct_Patm %>%
  mutate(Patm_avg = (Patm_MSY + Patm_rig)/2)

ggplot(Oct_Patm, aes(x = datetime, y = Patm_MSY, color = "MSY")) +
  geom_point() + 
  geom_point(data = Oct_Patm, aes(x = datetime, y = Patm_rig, color = "rig")) + 
  geom_point(data = Oct_Patm, aes(x = datetime, y = Patm_avg, color = "avg")) + 
  theme_minimal()

ggplot(Oct_MSY_met, aes(x = datetime, y = Patm, color = "MSY")) +
  geom_point() + 
  geom_point(data = Oct_rig_met, aes(x = datetime, y = Patm, color = "rig")) + 
  theme_minimal()

```

```{r}

# interpolate atmospheric pressure to GO data time stamps

GO <- read.csv("MAY_lat_long_temp_sal_Tin_air_Working.csv")
GO$date_time_utc <- as.POSIXct(GO$date_time_utc, format= "%Y-%m-%d %H:%M:%S")

GO_Patm <- data.frame(approx(May_Patm$datetime, May_Patm$Patm_avg, xout = GO$date_time_utc, rule = 2, method = "linear"))

GO <- cbind(GO, GO_Patm)

ggplot(GO, aes(x = x, y = y, color = "Patm")) +
  geom_point() + 
  geom_point(data = GO, aes(x = date_time_utc, y = licor.press, color = "Plicor")) + 
  theme_minimal()

GO <- GO %>% 
  select(-x) %>%
  rename(Patm = y)  

#write.csv(GO, here('MAY_lat_long_temp_sal_Tin_air_Working_Patm.csv'), row.names = FALSE)

```

## October cruise

# Pro Oceanus pCO2
```{r}

# read Pro Oceanus pCO2 data
POceanus <- read.csv(here('20221012pco2.csv')) 

# delete first row
POceanus <- POceanus[-(1),]
POceanus <- POceanus[-(69135:69147),]

# delete rows before underway water flowing
POceanus <- POceanus[-(1:5351),] #noon

```
# combine date time info and set column types

```{r}

POceanus$date_time <- paste(POceanus$Year, "-" ,POceanus$Month, "-", POceanus$Day, "x", POceanus$Hour, ":", POceanus$Minute, ":", POceanus$Second)
POceanus <- POceanus %>% relocate(date_time, .before = "Zero.A.D")

POceanus$date_time <- gsub(" ", "", POceanus$date_time)
POceanus$date_time <- gsub("x", " ", POceanus$date_time)

POceanus$date_time <- as.POSIXct(POceanus$date_time, tz = "GMT", format="%Y-%m-%d %H:%M:%OS")

POceanus$CO2 <- as.double(POceanus$CO2)
POceanus$Humidity <- as.double(POceanus$Humidity)
POceanus$Cell.gas.pressure <- as.double(POceanus$Cell.gas.pressure)
POceanus$IRGA.temperature <- as.double(POceanus$IRGA.temperature)
POceanus$Day <- as.double(POceanus$Day)

```


# Plot Pro Oceanus data

```{r}

ggplot(POceanus, aes(x = date_time, y = CO2))+
  geom_point()+
  theme_minimal()

ggplot(POceanus, aes(x = date_time, y = Humidity))+
  geom_point()+
  theme_minimal()

ggplot(POceanus, aes(x = date_time, y = Cell.gas.pressure))+
  geom_point()+
  theme_minimal()

scale = 0.2    
ggplot(POceanus, aes(x = date_time, y = CO2)) +
  geom_point(aes(color = "pCO2 ppm")) +
  geom_point(aes(y = Cell.gas.pressure/scale, color = "Cell gas pressure")) +
  scale_y_continuous(sec.axis = sec_axis(~.*scale, name="pressure")) +
  labs(x = "Date and Time", y = "pCO2 ppm", color = "") +
  scale_color_manual(values = c("orange2", "gray30"))+
  theme_minimal()


ggplot(POceanus, aes(x = date_time, y = IRGA.temperature))+
  geom_point()+
  theme_minimal()

scale = 0.005
ggplot(POceanus, aes(x = date_time, y = CO2)) +
  geom_point(aes(color = "pCO2 ppm")) +
  geom_point(aes(y = Humidity/scale, color = "Humidity")) +
  #scale_x_continuous(breaks = seq(0, 336, 24)) +
  scale_y_continuous(sec.axis = sec_axis(~.*scale, name="Humidity (units)")) +
  labs(x = "Date and Time", y = "pCO2 ppm", color = "") +
  scale_color_manual(values = c("orange2", "gray30"))+
  theme_minimal()


```
# data cleaning steps

```{r}

outliers <- POceanus %>% filter(CO2 > 4000)

# two outliers with CO2 > 4000, 
# second at 2022-10-12 06:57:16 when instrument was started for the day # filter this one out
# first at 2022-10-11 19:29:13
# was this a zero? 
# after this, CO2 and pressure drop, and take a while to stabilize again (circa 1 hr)
# CO2 seems to return to where it was

# take a closer look at October 11
POceanus_11 <- POceanus %>% filter(Day < 12)
ggplot(POceanus_11, aes(x = date_time, y = CO2))+
  geom_point()+
  theme_minimal()

scale = 0.05    
ggplot(POceanus_11, aes(x = date_time, y = CO2)) +
  geom_point(aes(color = "pCO2 ppm")) +
  geom_point(aes(y = (Cell.gas.pressure-800)/scale, color = "Cell gas pressure")) +
  scale_y_continuous(sec.axis = sec_axis(~800+.*scale, breaks = seq(800, 1300, 50), name="pressure")) +
  labs(x = "Date and Time", y = "pCO2 ppm", color = "") +
  scale_color_manual(values = c("orange2", "gray30"))+
  theme_minimal()

# filter out points above 4k
POceanus_11_under4k <- POceanus_11 %>% filter(CO2 < 4000)
# rename columns to match GO data frame
POceanus_11_under4k <- rename(POceanus_11_under4k, date_time_gmt = date_time)
POceanus_11_under4k <- rename(POceanus_11_under4k, CO2_ppm = CO2)

# plot data below 4k
ggplot(POceanus_11_under4k, aes(x = date_time_gmt, y = CO2_ppm))+
  geom_point()+
  theme_minimal()

# filter out points above 4k
POceanus_under4k <- POceanus %>% filter(CO2 < 4000)
# rename columns to match GO data frame
POceanus_under4k <- rename(POceanus_under4k, date_time_gmt = date_time)
POceanus_under4k <- rename(POceanus_under4k, CO2_ppm = CO2)

# plot data below 4k
ggplot(POceanus_under4k, aes(x = date_time_gmt, y = CO2_ppm))+
  geom_point()+
  theme_minimal()


#TODO: clean up long equilibration period after 2022-10-11 19:29:13

```
# read in October GO data

```{r}
# read in October GO files
GOfiles <- dir(here("MSR_GO_data_Oct/"), "*.txt") # get file names

GO_Oct <- GOfiles %>% map_dfr(~ read_tsv(here("MSR_GO_data_Oct/", .), show_col_types = FALSE, ))

# handle date and time columns
GO_Oct <- rename(GO_Oct, PC_Date = "PC Date")
GO_Oct <- rename(GO_Oct, PC_Time = "PC Time")

# combine date and time
GO_Oct$PC_Date <- as.character(GO_Oct$PC_Date) 
GO_Oct$PC_Time <- as.character(GO_Oct$PC_Time) 
GO_Oct$date_time_gmt <- paste(GO_Oct$PC_Date, GO_Oct$PC_Time)
GO_Oct$date_time_gmt <- as.POSIXct(GO_Oct$date_time_gmt, tz = "GMT", format="%d/%m/%y %H:%M:%OS")

GO_Oct <- rename(GO_Oct, CO2_ppm = "CO2 um/m")

```
# Plot raw GO data

```{r}

ggplot(GO_Oct, aes(x = date_time_gmt, y = CO2_ppm))+
  geom_point()+
  theme_minimal()

```

```{r}

GO_Oct_EQU <- GO_Oct %>% filter(Type == "EQU" | Type == "EQU-DRAIN")

ggplot(GO_Oct_EQU, aes(x = date_time_gmt, y = CO2_ppm))+
  geom_point()+
  theme_minimal()

#TODO: clean up beginning and end of Oct 11

GO_Oct_ATM <- GO_Oct %>% filter(Type == "ATM" | Type == "ATM-DRAIN")

ggplot(GO_Oct_ATM, aes(x = date_time_gmt, y = CO2_ppm))+
  geom_point()+
  theme_minimal()

GO_Oct_STD <- GO_Oct %>% 
  filter(Type == "STD1" | Type == "STD1-DRAIN" | Type == "STD2" | Type == "STD2-DRAIN" | Type == "STD3" | Type == "STD3-DRAIN" | Type == "STD4" | Type == "STD4-DRAIN" )

ggplot(GO_Oct_STD, aes(x = date_time_gmt, y = CO2_ppm))+
  geom_point()+
  theme_minimal()

```

```{r}


ggplot(GO_Oct_EQU, aes(x= date_time_gmt, y = CO2_ppm, color = "General Oceanics"))+ geom_point() +geom_point(data = POceanus_under4k, aes(color = "Pro Oceanus"))

```
# Read in GPS data

```{r}

gps_oct <- read_delim("sync_gps_ctd_oct.txt", show_col_types = FALSE)

gps_oct$dist_km <- gps_oct$Distancem/1000

gps_oct <- gps_oct %>% relocate(dist_km, .after = "Distancem")

```



```{r}
# match gps data to GO data based on datetime

for (i in 1:nrow(GO_Oct_EQU)) {
  
  GO_Oct_EQU$ind[i] <- which.min(abs(GO_Oct_EQU$date_time_gmt[i] - gps_oct$DateGMT))
  
}

gps_oct$row_num <- seq.int(nrow(gps_oct))
gps_oct$bottlerow <- as.character(gps_oct$row_num)
GO_Oct_EQU$bottlerow <- as.character(GO_Oct_EQU$ind)

GO_Oct_EQU_gps <- left_join(GO_Oct_EQU, gps_oct, by = "bottlerow")

# mapping$row_num <- seq.int(nrow(mapping))
# mapping$bottlerow <- as.character(mapping$row_num)
# GO_May$bottlerow <- as.character(GO_May$ind)
# 
# GOplus <- left_join(GO_May, mapping, by = "bottlerow")

```

```{r}

ggplot(GO_Oct_EQU_gps, aes(x = Latitude, y = CO2_ppm))+
  geom_point()+
  theme_minimal()

ggplot(GO_Oct_EQU_gps, aes(x = dist_km, y = CO2_ppm))+
  geom_point()+
  theme_minimal()

```

```{r}

#TODO update to October GO data

# interpolate atmospheric pressure to GO data time stamps

# GO <- read.csv("MAY_lat_long_temp_sal_Tin_air_Working.csv")
# GO$date_time_utc <- as.POSIXct(GO$date_time_utc, format= "%Y-%m-%d %H:%M:%S")
# 
# GO_Patm <- data.frame(approx(May_Patm$datetime, May_Patm$Patm_avg, xout = GO$date_time_utc, rule = 2, method = "linear"))
# 
# GO <- cbind(GO, GO_Patm)
# 
# ggplot(GO, aes(x = x, y = y, color = "Patm")) +
#   geom_point() + 
#   geom_point(data = GO, aes(x = date_time_utc, y = licor.press, color = "Plicor")) + 
#   theme_minimal()
# 
# GO <- GO %>% 
#   select(-x) %>%
#   rename(Patm = y)  

#write.csv(GO, here('MAY_lat_long_temp_sal_Tin_air_Working_Patm.csv'), row.names = FALSE)

```
# Convert wind speed units and interpolate wind speed

```{r}


# Oct
Oct_MSY_wind <- data.frame(approx(Oct_MSY_met$datetime, Oct_MSY_met$speed_mph, xout = Oct_rig_met$datetime, rule = 2, method = "linear"))

Oct_rig_wind <- Oct_rig_met %>% select(datetime, speed_mph)

Oct_wind <- cbind(Oct_MSY_wind, Oct_rig_wind)
Oct_wind <- Oct_wind %>% 
  rename(speed_mph_MSY = y) %>% 
  rename(speed_mph_rig = speed_mph) 

Oct_wind <- Oct_wind %>%
  mutate(wind_avg_mph = (speed_mph_MSY + speed_mph_rig)/2)

# convert mph to m/s
Oct_wind$wind_avg_m_s <- Oct_wind$wind_avg_mph*0.44704

# interpolate to GO timestamps
GO_Oct_wind <- data.frame(approx(Oct_wind$datetime, Oct_wind$wind_avg_m_s, xout = GO_Oct$date_time_utc, rule = 2, method = "linear"))

GO_Oct_wind <- GO_Oct_wind %>%
  rename(date_time_utc = x) %>%
  rename(wind_avg_m_s = y)

# output csv
write.csv(GO_Oct_wind, "GO_Oct_wind.csv", row.names = FALSE)

# plot
ggplot(Oct_wind, aes(x = datetime, y = speed_mph_MSY, color = "MSY")) +
  labs(x = "datetime", y = "MSY wind mph") +
  geom_point() + 
  geom_point(data = Oct_wind, aes(x = datetime, y = speed_mph_rig, color = "rig")) + 
  labs(x = "datetime", y = "rig wind mph") +
  geom_line(data = Oct_wind, aes(x = datetime, y = wind_avg_mph, color = "avg")) + 
  labs(x = "datetime", y = "wind avg mph") +
  theme_minimal()

ggplot(Oct_wind, aes(x = datetime, y = wind_avg_m_s)) +
  geom_point() + 
  geom_line() +
  scale_color_manual(values = "darkred") + # this is not setting the color
  labs(x = "datetime", y = "Avg wind speed m/s") +
  theme_minimal()

```




