---
title: "mississippi-data-assembly"
author: "Kate Morkeski"
date: "`r Sys.Date()`"
output: html_document
---
## Setup

Libraries used

```{r}

library(here)
library(readxl)
library(tidyverse)

```
## May cruise

# Read in CTD and bottle data

```{r}
compiled <- read_excel(here('MSR_1_Compiled.xlsx')) 
#salinity <- read_excel(here('MSR_1_Salinity.xlsx'))
```

# Format date and time

```{r}

compiled <- compiled %>%
  rename(time_utc = "Time (UTC)") %>%
  rename(station = Station_ID) %>%
  rename(cast = Cast_number) %>%
  rename(niskin = Niskin_ID) %>%
  rename(lat = "Latitude [decimal degrees]") %>%
  rename(long = "Longitude [decimal degrees]") %>%
  rename(temperature = "CTD_Potential_Temp [deg_C]") %>%
  rename(depth_m = "Depth [m]")

# combine date and time
compiled$Date <- as.character(compiled$Date) 
compiled$time_utc <- as.character(compiled$time_utc)
compiled$time_utc <- gsub("1899-12-31 ", "", compiled$time_utc)
compiled$date_time_utc <- paste(compiled$Date, compiled$time_utc)
compiled$date_time_utc <- as.POSIXct(compiled$date_time_utc, tz = "GMT", format="%Y-%m-%d %H:%M:%OS")

compiled <- compiled %>% relocate(date_time_utc, .before = station)

```

# Correct apparent error in Station 6 coordinate

```{r}

# cast 6, niskin 6 entered with lat/long identical to station 7 entries
compiled <- compiled %>% 
  mutate(lat = replace(lat, station == 6 & niskin == 6, 29.35218)) %>%
  mutate(long = replace(long, station == 6 & niskin == 6, -89.46450))  
 

```

# Create map

```{r}

# very generic state boundary
# need more detailed map data
us_map <- map_data("state")

louisiana <- us_map %>% 
  filter(region %in% "louisiana") 

mouth <- louisiana %>% filter(long > -91.5 & lat < 30.5)

mapping <- compiled %>% 
  select(date_time_utc, station, cast, niskin, lat, long, temperature, depth_m, CTDSAL_PSS78) 
  
ggplot(louisiana, aes(x = long, y = lat)) +
  geom_path() +
  geom_point(data = mapping, aes(x = long, y = lat, color = CTDSAL_PSS78)) +  
  theme_minimal()

ggplot(louisiana, aes(x = long, y = lat)) +
  geom_path() +
  geom_point(data = mapping, aes(x = long, y = lat, color = station)) +  
  theme_minimal()  
  
ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_point(data = mapping, aes(x = long, y = lat, color = CTDSAL_PSS78)) +  
  theme_minimal()

ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_text(data = mapping, aes(x = long, y = lat, label = station)) +  
  theme_minimal()

ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_text(data = mapping, aes(x = long, y = lat, label = station), size = 3) +  
  coord_fixed(xlim = c(-90.0, -89.25), ylim = c(28.5, 29.25))+
  theme_minimal()

ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_path(data = mapping, aes(x = long, y = lat, color = CTDSAL_PSS78)) +  
  coord_fixed(xlim = c(-90.0, -89.25), ylim = c(28.5, 29.25))+
  theme_minimal()


```

# Label stations

```{r}
mapping <- mapping %>%
  mutate(leg = case_when(station <10 ~ "river",
                         station == 10 ~ "gulf",
                         station > 14 ~ "gulf",
                         TRUE ~ "spur"))
  
```


# Towards join with GO data

```{r}

GOfiles <- dir(here("MSR_GO_data_May/"), "*.txt") # get file names

GOdata <- GOfiles %>% map_dfr(~ read_tsv(here("MSR_GO_data_May/", .), show_col_types = FALSE, ))

```

```{r}

GOdata <- rename(GOdata, PC_Date = "PC Date")
GOdata <- rename(GOdata, PC_Time = "PC Time")

# combine date and time
GOdata$PC_Date <- as.character(GOdata$PC_Date) 
GOdata$PC_Time <- as.character(GOdata$PC_Time) 
GOdata$date_time_utc <- paste(GOdata$PC_Date, GOdata$PC_Time)
GOdata$date_time_utc <- as.POSIXct(GOdata$date_time_utc, tz = "GMT", format="%d/%m/%y %H:%M:%OS")

```
# Match bottle data by time

## TODO: use lat long to calculate cruise distance, and interpolate that way instead

```{r}

  # match bottle data to GO data based on datetime
  # get lat, long, T, S from bottle data
  
  for (i in 1:nrow(GOdata)) {
    
    GOdata$ind[i] <- which.min(abs(GOdata$date_time_utc[i] - mapping$date_time_utc))
   
  }

mapping$row_num <- seq.int(nrow(mapping))
mapping$bottlerow <- as.character(mapping$row_num)
GOdata$bottlerow <- as.character(GOdata$ind)

GOplus <- left_join(GOdata, mapping, by = "bottlerow")
                    

```

# check time difference between GO system and nearest CTD time
# Label records more than 1 hr from a station
# Isolate data from equilibrator 

```{r}

GOplus$date_time_num <- as.numeric(GOplus$date_time_utc.x)
GOplus$time_diff <- (GOplus$date_time_utc.x - GOplus$date_time_utc.y)
GOplus$time_diff_num <- abs(as.numeric(GOplus$time_diff))
GOplus$time_diff_min <- GOplus$time_diff_num/60
GOplus$time_diff_hrs <- GOplus$time_diff_num/3600

hist(GOplus$time_diff_hrs)

GOplus <- GOplus %>%
  mutate(date_time_utc.y = replace(date_time_utc.y, time_diff_hrs > 1, NA))  %>%
  mutate(lat = replace(lat, time_diff_hrs > 1, NA))  %>%
  mutate(long = replace(long, time_diff_hrs > 1, NA))  %>%
  mutate(temperature = replace(temperature, time_diff_hrs > 1, NA))  %>%
  mutate(CTDSAL_PSS78 = replace(CTDSAL_PSS78, time_diff_hrs > 1, NA))  %>%
  mutate(station = replace(station, time_diff_hrs > 1, NA)) %>%
  mutate(cast = replace(cast, time_diff_hrs > 1, NA)) %>%
  mutate(niskin = replace(niskin, time_diff_hrs > 1, NA)) %>%
  mutate(leg = replace(leg, time_diff_hrs > 1, "> 1 hr from station")) %>%
  mutate(leg = replace(leg, date_time_num < 1652200657, "transit"))
#Based on GO log sheets, river data begin after 2022-05-10 16:37:37, which is 1652200657 in numeric format

GOplus <- GOplus %>% rename(CO2_ppm = "CO2 um/m")

GO_EQU_plus <- GOplus %>% filter(Type == "EQU" | Type == "EQU-DRAIN")

```
# Plot all GO EQU data

```{r}

ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_point(data = GO_EQU_plus, aes(x = long, y = lat, color = CO2_ppm)) + 
    scale_color_gradient(low="orange", high="darkred") +
  theme_minimal()  

ggplot(GO_EQU_plus, aes(x = date_time_utc.x, y = CO2_ppm, color = leg))+
  geom_point()+
  scale_color_manual(values = c("black", "darkturquoise", "chartreuse3", "yellow", "gray50"))+
  theme_minimal()

ggplot(GO_EQU_plus, aes(x = date_time_utc.y, y = CO2_ppm, color = leg))+
  geom_point()+
  scale_color_manual(values = c("black", "darkturquoise", "chartreuse3", "yellow", "gray50"))+
  theme_minimal()


```

```{r}


GO_EQU_cruise <- GO_EQU_plus %>% filter(leg != "transit")

```

# Plot data with pre-river data excluded

```{r}

ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_point(data = GO_EQU_cruise, aes(x = long, y = lat, color = CO2_ppm)) + 
    scale_color_gradient(low="orange", high="darkred") +
  theme_minimal()  

ggplot(GO_EQU_cruise, aes(x = date_time_utc.x, y = CO2_ppm, color = leg))+
  geom_point()+
  scale_color_manual(values = c("gray50", "darkturquoise", "chartreuse3", "yellow"))+
  theme_minimal()

ggplot(GO_EQU_cruise, aes(x = date_time_utc.y, y = CO2_ppm, color = leg))+
  geom_point()+
  scale_color_manual(values = c("gray50", "darkturquoise", "chartreuse3", "yellow"))+
  theme_minimal()

ggplot(GO_EQU_cruise, aes(x = date_time_utc.x, y = temperature, color = leg))+
  geom_point()+
  scale_color_manual(values = c("gray50", "darkturquoise", "chartreuse3", "yellow"))+
  theme_minimal()

ggplot(GO_EQU_cruise, aes(x = date_time_utc.x, y = CTDSAL_PSS78, color = leg))+
  geom_point()+
  scale_color_manual(values = c("gray50", "darkturquoise", "chartreuse3", "yellow"))+
  theme_minimal()


```

# Export file for GO processing

```{r}

# GO_out <- GOplus %>% 
#   rename("CO2 um/m" = CO2_ppm) %>%
#   rename(date_time_utc = date_time_utc.x) %>%
#   select(-ind, -bottlerow, -date_time_utc.y, -row_num, -date_time_num, -time_diff, -time_diff_num, -time_diff_min, -time_diff_hrs)
# 
# write.csv(GO_out, here('GO_Mississippi_May.csv'), row.names = FALSE)
# 
# GO_out <- GO_out %>% filter(leg != "transit") %>%
#                     filter(leg != "> 1 hr from station")
# 
# write.csv(GO_out, here('GO_Mississippi_May_near_stations.csv'), row.names = FALSE)

```



```{r}

# combine station and cast later

GOplus <- GOplus %>% filter(leg != "transit") 
   
GOplus <- GOplus %>% 
  group_by(station, cast, niskin) %>%
  mutate(min_time_diff = min(time_diff_num)) %>%
  ungroup()

closest_niskin <- GOplus %>% 
  group_by(station, cast, niskin) %>%
  summarize(min(time_diff_num)) 

GOplus$nearest_niskin <- NA_integer_

GOplus <- GOplus %>%
  mutate(nearest_niskin = case_when(time_diff_num == min_time_diff ~ niskin))

GOplus <- GOplus %>% 
  rename(date_time_utc = date_time_utc.x) %>%
  select(-niskin, -ind, -bottlerow, -date_time_utc.y, -row_num, -date_time_num, -time_diff, -time_diff_num, -time_diff_min, -time_diff_hrs, -min_time_diff)

# remove CTD parameters for non-niskin locations
GOplus$date_time_utc <- replace(GOplus$date_time_utc, is.na(GOplus$nearest_niskin), NA)
GOplus$station <- replace(GOplus$station, is.na(GOplus$nearest_niskin), NA)
GOplus$cast <- replace(GOplus$cast, is.na(GOplus$nearest_niskin), NA)
GOplus$lat <- replace(GOplus$lat, is.na(GOplus$nearest_niskin), NA)
GOplus$long <- replace(GOplus$long, is.na(GOplus$nearest_niskin), NA)
GOplus$temperature <- replace(GOplus$temperature, is.na(GOplus$nearest_niskin), NA)
GOplus$depth_m <- replace(GOplus$depth_m, is.na(GOplus$nearest_niskin), NA)
GOplus$CTDSAL_PSS78 <- replace(GOplus$CTDSAL_PSS78, is.na(GOplus$nearest_niskin), NA)
GOplus$leg <- replace(GOplus$leg, is.na(GOplus$nearest_niskin), NA)

# Check distribution of depth
hist(GOplus$depth_m)

# check relationship of temperature with depth
ggplot(GOplus, aes(x = depth_m, y = temperature, color = leg))+
  geom_point()+
  theme_minimal()
# temperature more strongly controlled by habitat than by depth

GO_out <- GOplus %>% 
  rename("CO2 um/m" = CO2_ppm)

# replace NAs with -999 for Matlab
GO_out <- GO_out %>% mutate(station = case_when(is.na(station) ~ -999, TRUE ~ station)) %>% 
                      mutate(cast = case_when(is.na(cast) ~ -999, TRUE ~ cast)) %>% 
                      mutate(lat = case_when(is.na(lat) ~ -999, TRUE ~ lat)) %>% 
                      mutate(long = case_when(is.na(long) ~ -999, TRUE ~ long)) %>% 
                      mutate(temperature = case_when(is.na(temperature) ~ -999, TRUE ~ temperature)) %>% 
                      mutate(depth_m = case_when(is.na(depth_m) ~ -999, TRUE ~ depth_m)) %>% 
                      mutate(CTDSAL_PSS78 = case_when(is.na(CTDSAL_PSS78) ~ -999, TRUE ~ CTDSAL_PSS78) )

GO_out <- GO_out %>% 
  filter(Type != "SHUT DOWN") 

write.csv(GO_out, here('GO_Mississippi_May.csv'), row.names = FALSE)

```
```{r}
# read in data processed with pCO2 Sys

xCO2 <- read.csv(here('MSP_20220510_Working_inter_stds.csv'))

xCO2_EQU <- xCO2 %>% 
  filter(Type == "EQU") 
  
# convert -999 to NA  
 xCO2_EQU <- na_if(xCO2_EQU, "-999")
 
 # calculate difference between corrected and raw CO2
 xCO2_EQU <- xCO2_EQU %>%
   mutate(xdiff = xCO2.corr - CO2.um.m)


```

```{r}

ggplot(xCO2_EQU, aes(x = YDay.Calc, y = xCO2.corr, color = "corrected xCO2"))+
  geom_point()+
  geom_point(aes(x = YDay.Calc, y = CO2.um.m, color = "raw CO2 um/m")) +
  theme_minimal()

ggplot(xCO2_EQU, aes(x = YDay.Calc, y = xdiff)) + geom_point()+theme_minimal()

ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_point(data = xCO2_EQU, aes(x = long, y = lat, color = CO2.um.m)) + 
    scale_color_gradient(low="orange", high="darkred") +
  theme_minimal() 


ggplot(mouth, aes(x = long, y = lat)) +
  geom_path() +
  geom_point(data = xCO2_EQU, aes(x = long, y = lat, color = xCO2.corr)) + 
    scale_color_gradient(low="orange", high="darkred") +
  theme_minimal()

```

## October cruise

# Pro Oceanus pCO2
```{r}

# read Pro Oceanus pCO2 data
POceanus <- read.csv(here('20221012pco2.csv')) 

# delete first row
POceanus <- POceanus[-(1),]
POceanus <- POceanus[-(69135:69147),]

# delete rows before underway water flowing
POceanus <- POceanus[-(1:5351),] #noon

```
# combine date time info and set column types

```{r}

POceanus$date_time <- paste(POceanus$Year, "-" ,POceanus$Month, "-", POceanus$Day, "x", POceanus$Hour, ":", POceanus$Minute, ":", POceanus$Second)
POceanus <- POceanus %>% relocate(date_time, .before = "Zero.A.D")

POceanus$date_time <- gsub(" ", "", POceanus$date_time)
POceanus$date_time <- gsub("x", " ", POceanus$date_time)

POceanus$date_time <- as.POSIXct(POceanus$date_time, format="%Y-%m-%d %H:%M:%OS")

POceanus$CO2 <- as.double(POceanus$CO2)
POceanus$Humidity <- as.double(POceanus$Humidity)
POceanus$Cell.gas.pressure <- as.double(POceanus$Cell.gas.pressure)
POceanus$IRGA.temperature <- as.double(POceanus$IRGA.temperature)
POceanus$Day <- as.double(POceanus$Day)

```


# Plot Pro Oceanus data

```{r}

ggplot(POceanus, aes(x = date_time, y = CO2))+
  geom_point()+
  theme_minimal()

ggplot(POceanus, aes(x = date_time, y = Humidity))+
  geom_point()+
  theme_minimal()

ggplot(POceanus, aes(x = date_time, y = Cell.gas.pressure))+
  geom_point()+
  theme_minimal()

scale = 0.2    
ggplot(POceanus, aes(x = date_time, y = CO2)) +
  geom_point(aes(color = "pCO2 ppm")) +
  geom_point(aes(y = Cell.gas.pressure/scale, color = "Cell gas pressure")) +
  scale_y_continuous(sec.axis = sec_axis(~.*scale, name="pressure")) +
  labs(x = "Date and Time", y = "pCO2 ppm", color = "") +
  scale_color_manual(values = c("orange2", "gray30"))+
  theme_minimal()
  

ggplot(POceanus, aes(x = date_time, y = IRGA.temperature))+
  geom_point()+
  theme_minimal()

scale = 0.005
ggplot(POceanus, aes(x = date_time, y = CO2)) +
  geom_point(aes(color = "pCO2 ppm")) +
  geom_point(aes(y = Humidity/scale, color = "Humidity")) +
  #scale_x_continuous(breaks = seq(0, 336, 24)) +
  scale_y_continuous(sec.axis = sec_axis(~.*scale, name="Humidity (units)")) +
  labs(x = "Date and Time", y = "pCO2 ppm", color = "") +
  scale_color_manual(values = c("orange2", "gray30"))+
  theme_minimal()


```
# data cleaning steps

```{r}

outliers <- POceanus %>% filter(CO2 > 4000)

# two outliers with CO2 > 4000, 
  # second at 2022-10-12 06:57:16 when instrument was started for the day # filter this one out
  # first at 2022-10-11 19:29:13
        # was this a zero? 
        # after this, CO2 and pressure drop, and take a while to stabilize again (circa 1 hr)
        # CO2 seems to return to where it was

POceanus_11 <- POceanus %>% filter(Day < 12)
ggplot(POceanus_11, aes(x = date_time, y = CO2))+
  geom_point()+
  theme_minimal()

scale = 0.05    
ggplot(POceanus_11, aes(x = date_time, y = CO2)) +
  geom_point(aes(color = "pCO2 ppm")) +
  geom_point(aes(y = (Cell.gas.pressure-800)/scale, color = "Cell gas pressure")) +
  scale_y_continuous(sec.axis = sec_axis(~800+.*scale, breaks = seq(800, 1300, 50), name="pressure")) +
  labs(x = "Date and Time", y = "pCO2 ppm", color = "") +
  scale_color_manual(values = c("orange2", "gray30"))+
  theme_minimal()

# filter out points above 4k
POceanus_11_under4k <- POceanus_11 %>% filter(CO2 < 4000)
ggplot(POceanus_11_under4k, aes(x = date_time, y = CO2))+
  geom_point()+
  theme_minimal()

#TODO: clean up long equilibration period after 2022-10-11 19:29:13

```





